---
title: "Lab 01 Report"
author: "KNM"
date: "January 24, 2019"
output: html_document
---
```{r, include=F}
knitr::opts_chunk$set(fig.width=10) 
```

## Comparison of Alcohol Users and Non-users
In this exercise, we will attempt to provide an answer to the question, "How do light users of alcohol differ from those who are heavy users of alcohol in terms of their self-reported overall health?" In this document we will:

* load our data
* make a definition of "light users" and "heavy users"
* filter our data to create two new, smaller datasets: a "light users" dataset and a "heavy users" dataset
* combine those two datasets
* prepare a large variety of graphs displaying the distribution of self-reported overall health of both groups
* report the results
* give several possible conclusions based on these results

Our group will compare people who use alcohol to people who do not use alcohol.  We will compare the distribution of Non-users, Light Users, Moderate Users, and Heavy Users. Finally, we will  illustrate the difference of heavy alcohol users to light alcohol users.

Our central goal is to illustrate the frequency of alcohol use overall as well as illustrate how the severity of alcohol use differs within alcohol users. We also want to determine if heavy users of alcohol are less healthy than light users.

For the purposes of this data, we will define the following terms:

A Current User is a person who has used alcohol in the past 30 days.
A Former User is a person who has not used alcohol in the past 30 days.
A Non-user is a person who has not used alcohol in their lifetime.
A Light User is a person who has drank less than 10 days in the past month.
A Moderate User is a person who has drank between 10 and 20 days in the past month.
A Heavy User is a person who has drank over 20 days in the past month.
Health is a self-reported variable. 

## Data

Data is from the National Survey on Drug Use and Health , which is conducted annually by the Substance Abuse and Mental Health Services Administration (SAMHSA), an agency in the U.S. Department of Health and Human Services (DHHS).  More information about this survey can be found at [ICPSR's data archive page](http://www.icpsr.umich.edu/icpsrweb/ICPSR/series/64).  The data has been prepared into a .Rdata file, so we can simply load it in the workspace with the `load()` command. At this time, we'll also load our packages:

```{r data_and_packages_load, message=F}
library(tidyverse)
library(forcats)
load("36361-0001-Data.rda")
```

Because the data object has an unpleasant name, *da36361.0001*, we make a copy with the name *full_data* and remove the original data object.

```{r data_rename}
full_data <- da36361.0001
remove(da36361.0001)
```

# Definition of "Current Users"

#Now, we need to determine what we will consider to be a "current user."  We will base this definition off of the number of a days that a respondent reports using alcohol in the past 30 days.

```{r reg_use_def}
#current_use_criteria <- !0
#if(full_data$ALCREC <= 30){
#  current_use
#}
# never_use_criteria <- 0
```

# Definition of "Light User"

We have defined light users as those who have drank alcohol 10 days or less within the past month.

```{r light_user_def}
light_use_criteria <- 10
```

# Definition of "Heavy User"

We have defined heavy users as those who have drank alcohol more than 20 days within the past month.

```{r heavy_user_def}
heavy_use_criteria <- 20
```

## Data Manipulation

We ar looking for the self-reported overall health of people who are categorized as "heavy," "light," "moderate," or "never" users. Below we have manipulated the data to classify people under these definitions using the variable *alc_user_status*.

```{r full_data}
full_data <- mutate(full_data, alc_user_status =
                    ifelse(ALCFLAG == "(0) Never used (IRALCRC = 9)", "Never User", 
                           ifelse(ALCMON == "(0) Did not use in the past month (IRALCRC = 2-3,0)", 
                                  "Former User", "Current User")))
full_data <- mutate(full_data, alc_30day_use = cut(ALCDAYS, c(0, 10, 20, 30),
                                                   c("Light User", "Moderate User", "Heavy User")))

full_data <- mutate(full_data, alc_user_status = ifelse(alc_user_status =="Current User",
                                                        as.character(alc_30day_use), alc_user_status))

full_data <- mutate(full_data, alc_user_status = as.factor(alc_user_status))

full_data <- mutate(full_data, alc_user_status = fct_relevel(alc_user_status,
                                                            "Never User",
                                                            "Light User",
                                                            "Moderate User",
                                                            "Heavy User"))
```

Using these definitions, we can now create two data subgroups: heavy users and light users. Now, we wish to compare the relative health of people who we have defined as heavy users versus those we have defined as light users. 

```{r subsetting_data}
heavy_users <- filter(full_data, ALCDAYS > heavy_use_criteria)
light_users <- filter(full_data, ALCDAYS < light_use_criteria)
```

Next, we remove the *full_data* object and drop the unneeded variables from our *light_users* and *heavy_users* datasets.  These steps are not necessary, but clearing out unneeded objects and variables can improve the speed with with your code is executed.

```{r shrinking_data}
# remove(full_data)

heavy_users <- select(heavy_users, HEALTH)
light_users <- select(light_users, HEALTH)
```

It may be useful to us to have the data about both groups (heavy and light users) in our data object.  To accomplish this, we add a new variable *user_status* to both our data objects.  For all observations in *current_users* we wish to set the value of this variable to "current"  For all observations in *light_users* we wish to set the value of this variable to "light"  Once we have added this variable to both datasets, we combine them using the **rbind()** function.  **rbind()** stands for row bind -- it binds or glues two datasets together along by putting one dataset on top of the other.

```{r combining data}
heavy_users$alc_user_status <- "heavy"
light_users$alc_user_status <- "light"
combined_data <- rbind(heavy_users, light_users)
```

# Visual Displays

We start our graphical analysis by making a some basic plots showing the self-reported overall health of regular users.

```{r display}
ggplot(full_data)+
  geom_bar(aes(x=alc_user_status, fill=HEALTH),position="fill")+
  scale_x_discrete(name="Use of Alcohol",labels=c("Never User",
                            "Former User",
                            "Light Current User\n(10 or fewer days per month)",
                            "Moderate Current User\n(11-20 days per month)",
                            "Heavy Current User\n(more than 20 days per month)"))+
  scale_fill_brewer(palette="RdYlGn",direction=-1,
                    labels=c("Excellent","Very Good","Good","Fair","Poor"),
                    name="Self-Reported \nOverall Health")+
  coord_flip()+labs(y="Proportion",title="Self-Reported Health by Marijuana User Status")+
  theme_classic()
```
## Comparison of Non-Alcohol Users and Alcohol Users Based on Severity

```{r bar_graph_fulldata}
ggplot(full_data)+geom_bar(aes(x=alc_user_status, fill=HEALTH),position="fill")+scale_fill_brewer(palette="RdYlGn", direction=-1, labels=c("Excellent", "Very Good", "Good", "Fair", "Poor"), name="Self-Reported \nOverall Health") + labs(x="User Status", y="Relative Frequency", title="Self-Reported Health by Alcohol User Status")+theme_classic() 
```

This graph displays the self-reported overall health of people who are categorized as "never," "light," "moderate," or "heavy" users.

Now, we wish to compare the relative health of people who we have defined as heavy users versus those we have defined as light users. We defined heavy users as those who have drank alcohol 30 days within the past month (30 days) and light users as those who have drank 10 days within the past month.

## Bar Graphs Comparing Heavy and Heavy Users
```{r bar_plot_combined_data}
ggplot(combined_data)+
  geom_bar(aes(x=HEALTH, fill=HEALTH, y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]))+scale_fill_brewer(palette="RdYlGn", direction=-1,labels=c("Excellent", "Very Good", "Good", "Fair", "Poor"), name="Self-Reported \nOverall Health")+
  labs(x="", y="Relative Frequency", title="Self-Reported Health by User Status")+scale_x_discrete(labels=c("Excellent", "Very Good", "Good", "Fair", "Poor"))+facet_wrap("alc_user_status")+theme_classic()
```


## Pie Chart Comparing Heavy and Light Users
```{r combined_data_piegraph}
ggplot(combined_data) + geom_bar(aes(x=0, fill=HEALTH, y=..count../tapply(..count..,..PANEL..,sum)[..PANEL..]))+ coord_polar(theta="y") + scale_fill_brewer(palette="RdYlGn", direction =-1, labels=c("Excellent", "Very Good", "Good", "Fair", "Poor"), name="Self-Reported \nOverall Health")+ 
labs(x="", y="Relative Frequency", title="Self-Reported Health by User Status") +facet_wrap("alc_user_status")+theme_classic()
```

## Stacked Bar Plot Comparison of Heavy and Light Users
```{r stackedplot}
ggplot(combined_data)+geom_bar(aes(x=alc_user_status, fill=HEALTH),position="fill")+scale_fill_brewer(palette="RdYlGn", direction=-1, labels=c("Excellent", "Very Good", "Good", "Fair", "Poor"), name="Self-Reported \nOverall Health") + labs(x="User Status", y="Relative Frequency", title="Self Reported Health by User Status")+scale_x_discrete(labels=c("Heavy Users", "Light Users"))+theme_classic()
```


# Numerical Verifcation

```{r verification}
user_counts <- count(combined_data,alc_user_status)
health_counts <- count(combined_data,alc_user_status,HEALTH)
print(user_counts)
print(health_counts,n=nrow(health_counts))
```

Manually, we see that `r filter(health_counts,alc_user_status=="heavy",HEALTH=="(1) Excellent")[,"n"]` of `r user_counts[1,"n"]` (about `r round(filter(health_counts,alc_user_status=="heavy",HEALTH=="(1) Excellent")[,"n"]/user_counts[1,"n"]*100)`%) "heavy users" self-reported "excellent" overall health, which is consistent with what our display shows.

Test: `r user_counts[1,"n"]`

# Poor health and heavy

Manually, we see that `r filter(health_counts,alc_user_status=="heavy",HEALTH=="(5) Poor")[,"n"]` of `r user_counts[1,"n"]` (about `r round(filter(health_counts,alc_user_status=="heavy",HEALTH=="(5) Poor")[,"n"]/user_counts[1,"n"]*100)`%) "heavy users" self-reported "poor" overall health, which is consistent with what our display shows.

Manually, we see that `r filter(health_counts,alc_user_status=="light",HEALTH=="(5) Poor")[,"n"]` of `r user_counts[2,"n"]` (about `r round(filter(health_counts,alc_user_status=="light",HEALTH=="(5) Poor")[,"n"]/user_counts[2,"n"]*100)`%) "light users" self-reported "poor" overall health, which is consistent with what our display shows.

# Fair health

Manually, we see that `r filter(health_counts,alc_user_status=="heavy",HEALTH=="(4) Fair")[,"n"]` of `r user_counts[1,"n"]` (about `r round(filter(health_counts,alc_user_status=="heavy",HEALTH=="(4) Fair")[,"n"]/user_counts[1,"n"]*100)`%) "heavy users" self-reported "fair" overall health, which is consistent with what our display shows.

Manually, we see that `r filter(health_counts,alc_user_status=="light",HEALTH=="(4) Fair")[,"n"]` of `r user_counts[2,"n"]` (about `r round(filter(health_counts,alc_user_status=="light",HEALTH=="(4) Fair")[,"n"]/user_counts[2,"n"]*100)`%) "light users" self-reported "fair" overall health, which is consistent with what our display shows.

# Good health

Manually, we see that `r filter(health_counts,alc_user_status=="heavy",HEALTH=="(3) Good")[,"n"]` of `r user_counts[1,"n"]` (about `r round(filter(health_counts,alc_user_status=="heavy",HEALTH=="(3) Good")[,"n"]/user_counts[1,"n"]*100)`%) "heavy users" self-reported "good" overall health, which is consistent with what our display shows.

Manually, we see that `r filter(health_counts,alc_user_status=="light",HEALTH=="(3) Good")[,"n"]` of `r user_counts[2,"n"]` (about `r round(filter(health_counts,alc_user_status=="light",HEALTH=="(3) Good")[,"n"]/user_counts[2,"n"]*100)`%) "light users" self-reported "good" overall health, which is consistent with what our display shows.

# Very good health

Manually, we see that `r filter(health_counts,alc_user_status=="heavy",HEALTH=="(2) Very good")[,"n"]` of `r user_counts[1,"n"]` (about `r round(filter(health_counts,alc_user_status=="heavy",HEALTH=="(2) Very good")[,"n"]/user_counts[1,"n"]*100)`%) "heavy users" self-reported "very good" overall health, which is consistent with what our display shows.

Manually, we see that `r filter(health_counts,alc_user_status=="light",HEALTH=="(2) Very good")[,"n"]` of `r user_counts[2,"n"]` (about `r round(filter(health_counts,alc_user_status=="light",HEALTH=="(2) Very good")[,"n"]/user_counts[2,"n"]*100)`%) "light users" self-reported "very good" overall health, which is consistent with what our display shows.

# Excellent health

Manually, we see that `r filter(health_counts,alc_user_status=="heavy",HEALTH=="(1) Excellent")[,"n"]` of `r user_counts[1,"n"]` (about `r round(filter(health_counts,alc_user_status=="heavy",HEALTH=="(1) Excellent")[,"n"]/user_counts[1,"n"]*100)`%) "heavy users" self-reported "excellent" overall health, which is consistent with what our display shows.

Manually, we see that `r filter(health_counts,alc_user_status=="light",HEALTH=="(1) Excellent")[,"n"]` of `r user_counts[2,"n"]` (about `r round(filter(health_counts,alc_user_status=="light",HEALTH=="(1) Excellent")[,"n"]/user_counts[2,"n"]*100)`%) "light users" self-reported "excellent" overall health, which is consistent with what our display shows.

# Results

# Conclusion/Discussion