---
title: "In Class Work February 5th"
author: "Greg Wills"
date: "2/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction and Data
```{r load_data, include=FALSE}
# loading packages
library(tidyverse)
library(lubridate)
library(broom)
library(rgeos)
library(rgdal)
library(maptools)
library(mapproj)

# loading data files
sf_data <- read_csv("sf.csv")
sf_data_2017 <- read_csv("sf2017.csv")

#data cleanup
sf_data <- mutate(sf_data, date = mdy(Report_taken_date_EST),
                  Type=as.factor(`Data Type`),
                  Gender=as.factor(Subject_Sex),
                  Ethnicity=as.factor(Subject_Ethnicity),
                  Race=as.factor(Subject_Race),
                  District=as.factor(`Incident Location District`),
                  PSA=as.factor(`Incident Location PSA`),
                  Year=as.factor(Year)
)

# if age is juvenile or unknown = NA, otherwise age as numerical value
sf_data <- mutate(sf_data, num_age = 
                    ifelse(Age=="Juvenile"|Age=="Unknown",
                           NA,as.numeric(Age)))

# cutting age into 10-year groups
sf_data <- mutate(sf_data, cat_age = cut(num_age, 
                                         breaks=c(17,27,37,47,57,67,77,87),
                                         labels=c("18-27","28-37","38-47",
                                                  "48-57","58-67","68-77",
                                                  "78-87")))

# add "Juvenile" and "Unknown" into categorical age variable
sf_data <- mutate(sf_data, cat_age = ifelse(Age=="Juvenile"|Age=="Unknown",Age,
                                            as.character(cat_age)))

# rename categorical age variable to Age_binned, coerce to factor
# Make "Juvenile" the first level of that factor
sf_data <- mutate(sf_data, Age_binned=as.factor(cat_age))
sf_data <- mutate(sf_data, Age=fct_relevel(Age,"Juvenile"),
                  Age_binned=fct_relevel(Age_binned,"Juvenile"))

# create Month variable
sf_data <- mutate(sf_data, Month=month(date,label=T,abbr=F))

# import geospatial data about Police Service Areas
psas <- readOGR("Police_Service_Areas.shp",layer="Police_Service_Areas")

# extracting the lat/long coordinates of PSA boundary points
# telling the tidy function that the region value can be found in the "PSA" variable
psas.points <- tidy(psas, region="PSA")

# combine lat/long coordinates of district boundary points with data about districts
# tell join function that "id" in psas.points corresponds to "PSA" in psas@data
psas.cart <- full_join(psas.points, psas@data, by=c("id" = "PSA"))
```

# Graph 1
```{r graph1}

plot_data <- filter(sf_data,!is.na(PSA))

all_cats <- expand(plot_data,PSA,Gender)

plot_data <- summarize(group_by(plot_data,PSA,Gender),reports=n())

plot_data <- full_join(plot_data,all_cats,by=c("PSA","Gender"))

plot_data <- mutate(plot_data,reports=replace_na(reports,0))

names(plot_data)[1] <- "NAME"

mapdata <- full_join(plot_data,psas.cart,by="NAME")

g <- ggplot() +
  geom_polygon(data=drop_na(mapdata), aes(x=long, y=lat, group=group,fill=reports),color="black") +
  coord_map()+
  theme_void()
g+facet_wrap(~Gender)+
  scale_fill_distiller(palette="Spectral", name="Number of Reports")
```

Graph 1: A panel of maps showing the number of stop and frisk reports in each police service area faceted by gender

#Graph 2
```{r graph2}
# import geospatial data about Police Districts
districts <- readOGR("Police_Districts.shp",layer="Police_Districts")

# defining districts as "id" variable
names(districts@data)[1] <- "id"

# extracting the lat/long coordinates of district boundary points
# telling the tidy function that the region value can be found in the "id" variable
districts.points <- tidy(districts, region="id")

# combine lat/long coordinates of district boundary points with data about districts
districts.cart <- full_join(districts.points, districts@data, by="id")

plot_data2 <- filter(sf_data,!is.na(District))

levels(plot_data2$District) <- c("First District","Second District",
                                "Third District","Fourth District",
                                "Fifth District","Sixth District",
                                "Seventh District")

all_cats <- expand(plot_data2,District,Age_binned)

plot_data2 <- summarize(group_by(plot_data2, District,Age_binned),reports=n())

plot_data2 <- full_join(plot_data2,all_cats,by=c("District","Age_binned"))

#plot_data <- mutate(plot_data,reports=replace_na(reports,0))

names(plot_data2)[1] <- "NAME"

mapdata <- full_join(plot_data2,districts.cart,by="NAME")

g <- ggplot() +
  geom_polygon(data=mapdata,
               aes(x=long, y=lat, group=group,fill=reports),color="black") +
  coord_map()+
  theme_void()

g+facet_wrap(~Age_binned)+
  scale_fill_distiller(palette="Spectral", name="Number of Reports",na.value="white")
```

Graph 2: A panel of maps showing the number of stop and frisk reports in each police district faceted by age group

#Graph 3
```{r graph3}
crime <- read_csv("CrimeStatebyState.csv")
us_data <- filter(crime,State == "United States-Total")
ggplot(us_data, aes(x=Year, y=`Violent Crime rate`)) +
  geom_line(color="green") + 
  labs(title="Violent Crime in the United States", y="Crimes per 100,000 residents") +
  theme_classic()
```

Graph 3: A time series plot showing the violent crime rate in the United States overall from 1960 to 2014 as a green line

#Graph 4
```{r graph4}
crime <- read_csv("CrimeStatebyState.csv")
# 
#crime_filtered <- summarize(group_by(crime, Year,`Aggravated assault rate`,`Murder and nonnegligent manslaughter rate`,`Legacy rape rate /1`,`Robbery rate`))

us_crimes <- select(crime, Year,`Aggravated assault rate`,`Murder and nonnegligent manslaughter rate`,`Legacy rape rate /1`,`Robbery rate`)

us_crimes_gathered <- gather(us_crimes, key="Violent Crime Type",value=Rate,-Year)

ggplot(us_crimes_gathered) + 
  geom_line(aes(x=Year,y=Rate,color=`Violent Crime Type`))+
  scale_color_discrete(name="Violent Crime Type")+
  theme_classic()

```

Graph 4: A time series plot showing the national rates for each of 4 violent crime categories (assault, murder/manslaughter, rape, and robbery) for years from 1960 to 2014 – a different color is used for each type of crime, include a legend mapping colors to crimes, labeled the horizontal axis “Year,” labeled the vertical axis “Rate”

#Graph 5
```{r graph5}
crime <- read_csv("CrimeStatebyState.csv")

state_crime<- filter(crime,State=="United States-Total")
colnames(crime)[colnames(crime)=="Property crime total"] <- "PropertyCrimeTotal"
colnames(crime)[colnames(crime)=="Violent crime total"] <- "ViolentCrimeTotal"
colnames(crime)[colnames(crime)=="Murder and nonnegligent Manslaughter"] <- "MurderTotal"
colnames(crime)[colnames(crime)=="Legacy rape /1"] <- "RapeTotal"
colnames(crime)[colnames(crime)=="Aggravated assault"] <- "AssaultTotal"
colnames(crime)[colnames(crime)=="Robbery"] <- "RobberyTotal"

northeast_states<- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania")
midwest_states <- c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota")
south_states <- c("Delaware", "Florida", "Georgia", "Maryland", "North Carolina", "South Carolina", "Virginia", "District of Columbia", "West Virginia", "Alabama", "Kentucky", "Mississippi", "Tennessee", "Arkansas", "Louisiana", "Oklahoma", "Texas")
west_states <- c("Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Utah", "Wyoming", "Alaska", "California", "Hawaii", "Oregon", "Washington")

state_crime <- mutate(crime, region=ifelse(State %in% northeast_states, "Northeast",
                                                 ifelse(State %in% midwest_states,"Midwest",
                                                        ifelse(State %in% south_states, "South",
                                                               ifelse(State %in% west_states, "West","NA")))))

regional_rates <- summarize(group_by(state_crime,region,Year),
                            property_rate=sum(PropertyCrimeTotal)/sum(Population)*100000,
                            violent_rate=sum(ViolentCrimeTotal)/sum(Population)*100000, 
                            murder_rate=sum(MurderTotal)/sum(Population)*100000,
                            rape_rate=sum(RapeTotal)/sum(Population)*100000 ,
                            assault_rate=sum(AssaultTotal)/sum(Population)*100000,
                            robbery_rate=sum(RobberyTotal)/sum(Population)*100000)


ggplot(regional_rates)+
  geom_line(aes(x=Year, y=property_rate, group=region, color=region))+
  xlim(1960, 2020)+
  labs(title="Overall property crime rate for each of the 4 census regions", y="Property Crime Rate\n")+
  theme_classic()
```

Graph 5: A time series plot showing the overall property crime rate for each of the 4 census regions (see the prior notes for R commands defining these regions and creating a variable designating them in this data). Note: After creating your region variable, you will need to use that variable to group your data and generate summary statistics; you will need to generate the summary statistics that will allow you to calculate the property crime rate for each state. Note that this region rate cannot be found using only the property crime rates of each state in the region.

#Graph 6
```{r graph6}
crime_2013 <- filter(crime,Year==2013, State!="United States-Total")
all_states <- map_data("state")
crime_2013 <- rename(crime_2013,region=State)
crime_2013 <- mutate(crime_2013, region=tolower(region))
stateData <- left_join(all_states,crime_2013,by="region")
ggplot()+
  geom_polygon(data=stateData,
               aes(x=long, y=lat, group = group, fill=`Robbery rate`),
               color="grey50")+
  scale_fill_gradient2(low="orange",high="darkorange",mid="white",midpoint=0)+
  coord_map() +
  labs(title="Robbery Rate in 2013", y="Latitude", x="Longitude")
```

Graph 6: A choropleth showing the robbery rate in 2013 in each state as shades of orange (darker orange indicating a high robbery rate and lighter orange indicating a lower robbery rate).

#Graph 7
```{r graph7}
library(treemap)
crime <- read_csv("CrimeStatebyState.csv")
crime_2010 <- filter(crime,Year==2010,State!="United States-Total")
crime_2010 <- rename(crime_2010,region=State)
crime_2010 <- mutate(crime_2010, region=tolower(region))
stateData <- left_join(all_states,crime_2010,by="region")

treemap(stateData,
        index="region",
        vSize="Population",
        vColor="Murder and nonnegligent Manslaughter",
        type="value",palette="Reds")
```

Graph 7: A cartogram in which the size of each rectangle is proportional to the population of the state depicting the murder rate in each state in 2010 as shades of red (darker red for higher rates)

#Graph 8
```{r graph8}
crime_1990 <- filter(crime,Year==1990, State!="United States-Total")

northeast_states<- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania")
midwest_states <- c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota")
south_states <- c("Delaware", "Florida", "Georgia", "Maryland", "North Carolina", "South Carolina", "Virginia", "District of Columbia", "West Virginia", "Alabama", "Kentucky", "Mississippi", "Tennessee", "Arkansas", "Louisiana", "Oklahoma", "Texas")
west_states <- c("Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Utah", "Wyoming", "Alaska", "California", "Hawaii", "Oregon", "Washington")

crime_1990 <- mutate(crime_1990, region=ifelse(State %in% northeast_states, "Northeast",
                                                 ifelse(State %in% midwest_states, "Midwest",
                                                        ifelse(State %in% south_states, "South",
                                                               ifelse(State %in% west_states, "West","NA")))))

ggplot(crime_1990)+
  geom_point(aes(x=`Motor vehicle theft rate`,
                 y=`Burglary rate`,
                 color=region,
                 shape=region))+
  scale_shape_manual(values=c("Midwest"=17,"Northeast"=18,"South"=15,"West"=19))+
  scale_color_manual(values=c("West"="red", "Northeast"="blue", "South"="green", "Midwest"="purple"))+
  labs(x="Motor Vehicle Theft Rate",y="Burglary Rate",title="Motor Vehicle and Burglary Rate\n by US Region")
```

Graph 8: A scatter plot showing (for each state) the motor vehicle theft rate in 1990 along the horizontal axis and the burglary rate in 1990 along the vertical axis. A different color and shape is used for the points representing states from each of the 4 regions. A solid green square is used for the South, a solid red circle for the West, a solid purple triangle for the Midwest, and a solid blue diamond for the Northest.

#Graph 9
```{r graph9}
crime <- read_csv("CrimeStatebyState.csv")
crime_2014 <- filter(crime,Year==2014,State!="United States-Total")

northeast_states<- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania")
midwest_states <- c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota")
south_states <- c("Delaware", "Florida", "Georgia", "Maryland", "North Carolina", "South Carolina", "Virginia", "District of Columbia", "West Virginia", "Alabama", "Kentucky", "Mississippi", "Tennessee", "Arkansas", "Louisiana", "Oklahoma", "Texas")
west_states <- c("Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Utah", "Wyoming", "Alaska", "California", "Hawaii", "Oregon", "Washington")

crime_2014 <- mutate(crime_2014, region=ifelse(State %in% northeast_states, "Northeast",
                                                 ifelse(State %in% midwest_states, "Midwest",
                                                        ifelse(State %in% south_states, "South",
                                                               ifelse(State %in% west_states, "West","NA")))))

ggplot(crime_2014)+
  geom_point(aes(x=`Larceny-theft rate`,y=`Aggravated assault rate`,size=`Population`,color=region))+
  labs(x="Larceny Theft Rate",y="Aggravated Assault Rate",title="Relationship between larceny theft rate rate and aggravated assault rate in 2014\n (Size of bubbles indicates population size)")
```

Graph 9: A scatter plot showing (for each state) the larceny theft rate rate in 2014 along the horizontal axis and the aggravated assault rate in 2014 in along the vertical axis. The size of each point is mapped to that state’s population.